---
title: "Untitled"
author: "Marc Weber"
date: "5 1 2022"
output: html_document
---
```{r}
library(nnet)
library(ROCR)
```


# Neural Networks

## Build the network

```{r echo=TRUE, error=FALSE, message=FALSE, warning=FALSE, include=TRUE}
set.seed(1400)

# Work with copy of df.cc
df.ann <- cbind(df.cc)

# Standardize
df.ann$log_amt_income_total <- log(df.ann$amt_income_total)

## Only get needed columns for model
keeps <- c(
          "customer_status_good"
          ,"code_gender"
          ,"age_years"
          ,"name_education_type"
          ,"name_income_type"
          ,"name_housing_type"
          ,"name_family_status"
           ,"cnt_fam_members"
          ,"flag_own_car"
          ,"flag_own_realty"
          , "log_amt_income_total"
          ,"occupation_type"
           )

df.ann <- df.ann[keeps]


# Split data into train and test data sets
intrain <- createDataPartition(y = df.ann$customer_status_good,
                               p=0.65,
                               list = FALSE)
df.ann.training <- df.ann[intrain,]
df.ann.testing <- df.ann[-intrain,]

# Fit model
df.ann.net <- nnet(customer_status_good ~ ., data = df.ann.training, size=15, maxit=500, range=0.1, decay=5e-4, MaxNWts = 10000, trace = FALSE)
```

The model is designed as 41-15-1 network with 646 weights. 11 predictors were added to the model, but it works in total with 41 predictors as some of the variables are factors, which will be converted to dummy variables (0/1 values).



## Make Predictions


```{r}
pred <- predict(df.ann.net, df.ann.testing, type="class")
cm_nn <- table(pred=pred, true=df.ann.testing$customer_status_good)
cm_nn
```

Calculating the accuracy

```{r}
sum(diag(cm_nn))/sum(sum(cm_nn))
```

Percentage of identified 'bad' customers

```{r}
cm_nn[1, 1]/(cm_nn[1, 1] + cm_nn[2, 1])
```

```{r}
cm_nn[1, 1] + cm_nn[2, 1]
```


The accuracy indicates a good score (`r sum(diag(cm_nn))/sum(sum(cm_nn))`). However, simply predict all customers as good customers would lead to a even higher accurary (approx. 0.84). The model has issues to identify bad customers. It identifies `r cm_nn[1, 1] + cm_nn[1, 2]` bad customers, but most of them are in fact good customers. Out of the `r cm_nn[1, 1] + cm_nn[2, 1]` 'real' bad customers  the model identifies only `r cm_nn[1, 1]`.


```{r echo=TRUE, error=FALSE, message=FALSE, warning=FALSE, include=TRUE}
# Tune nnet with function for different size of neurons
x <- c()
for (i in 1:25) {
  df.ann.net <- nnet(customer_status_good ~ ., data = df.ann.training, size=i, maxit=100, range=0.1, decay=5e-4, MaxNWts = 10000, trace = FALSE)
  pred <- predict(df.ann.net, df.ann.testing, type="class")
  cm_nn <- table(pred=pred, true=df.ann.testing$customer_status_good)
  s <- sum(diag(cm_nn))/sum(sum(cm_nn))
  x <- append(x, s)
}
plot(x, main="Tune nnet - Count of Neurons", ylab= "Accuracy", xlab = "Count of Neurons")


```

This function helps to tune the amount of neurons for the model. However, there is no model, which helps to find a high accuracy and is able to identify bad customers well.

Importance of the predictors
```{r}
df.ann.net.imp <- varImp(df.ann.net)
df.ann.net.imp <- df.ann.net.imp[order(-df.ann.net.imp$Overall),, drop = FALSE]
head(df.ann.net.imp, 10)

```

This table shows the importance of the different predictors (list of top 10 out of 41). This output should not be overinterpreted as the model itself is not suitable to identify bad customers.

# Optimization
The distinction between good and bad customer is a way to build up a relatively safe portfolio of customers, who repay their debts with the interest. However, the defined threshold leaves room for optimization.  
  
Potential options to increase earnings through optimization include:  
  
<b>1. Granting credit cards to a percentage of "bad" applicants</b> <br>
This adds a marginally higher risk to the portfolio of credit card customers, however, comes with higher possible returns.  
  
<b> 2. Allocating credit limits  </b> <br>
Determining the credit limit to allow an applicant/customer based on their likelihood to repay (i.e. "percent_good") in order to maximize interest paid, thereby increasing earnings.  
  
<b> 3. Dunning</b> <br>
It is lucrative for credit card companies when customers do not immediately repay the loan amounts they borrow. Depending on the agreed contract, the companies can claim additional fees and penalty interest from a certain point. The dunning system should take this into account and not dun the customer too early or the full amount as overdue. However, it should be noted that the later the reminder is issued, the greater the likelihood of a credit default (untested assumption). It is therefore appropriate to demand regular partial payments.

To optimize (maximize) the profit of a credit card company the following <b>simplified</b> formular could be applied:

<b>Maximize $g(x,y) =$ “loan amount” $\cdot$ “interest rate” $\cdot \frac{x}{360} +$ “additional fees” $–$ “loan amount” $\cdot$ “probability of default”$(x,y)$</b>

Control variables: <br>
$x =$ days until dunning letter will be sent<br>
$y =$ amount of partial payment

Probability of default is another function based on $x$ and $y$. The higher the value of $x$ and the lower the value of $y$, it can be assumed that the probability of default will be increased.

Constraints are defined based on the agreed contracts between the company and the customers. For example, there it can be defined the minimum days until an amount is overdue (e.g. $x >=$ 30 days).


