---
title: "Machine learning models and credic card data"
author:
  - Students - 
  - Professors - Mateo
date: "Last edited `r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false 
      smooth_scroll: false
    toc_depth: 5
    number_section: true 
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
}
code.r{
  font-size: 11px;
}
pre {
  font-size: 10px
}
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 32px;
  font-weight: bold;
  text-align: center;
  color: #2A5A7A;
  opacity: 0.8;
}
h4.author { 
  font-size: 15px;
  color: #468faf;
  text-align: center;
}
h4.date { 
  font-size: 16px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h1 { /* Header 1 */
  font-size: 22px;
  color: Black;
  font-weight: bold;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: Black;
  font-weight: bold
}
h3 { /* Header 3 */
  font-size: 15px;
  color: Black;
  font-weight: bold
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(ggplot2)
library(tseries)               # for ts object, ADF test
library(magrittr)
library(data.table)
library(dplyr)
library(lubridate)

```

# Introduction

# Data preparation


```{r preparation1, include=FALSE}

# load data
df_app <- fread("data/application_record.csv", stringsAsFactors = TRUE)
df_cc <- fread("data/credit_record_pivot.csv", stringsAsFactors = TRUE)

```

```{r preparation2, include=FALSE}

# transform data
df_joined <- inner_join(x = df_app, y = df_cc, by = "ID") %>%
  mutate(ID = as.factor(ID),
          CNT_CHILDREN = as.factor(CNT_CHILDREN),
          CNT_FAM_MEMBERS = as.factor(CNT_FAM_MEMBERS),
          age_years = as.numeric(round(DAYS_BIRTH * (-1) / 365.25), 0),        # instead of DAYS_BIRTH
          work_years = as.numeric(round((DAYS_EMPLOYED * (-1) / 365.25), 0)),  # instead of DAYS_EMPLOYED,
                                                                              # negative: Pensioner or                                                                                                     unemployed
          FLAG_MOBIL = as.factor(FLAG_MOBIL),
          FLAG_WORK_PHONE = as.factor(FLAG_WORK_PHONE),
          FLAG_PHONE = as.factor(FLAG_PHONE),
          FLAG_EMAIL = as.factor(FLAG_EMAIL),
          past_0_29 = as.numeric(past_0_29),
          past_30_59 = as.numeric(past_30_59),
          past_60_89 = as.numeric(past_60_89),
          past_90_119 = as.numeric(past_90_119),
          past_120_149 = as.numeric(past_120_149),
          past_150_plus = as.numeric(past_150_plus)
  )

# Change column names to lower case
setnames(df_joined, tolower(names(df_joined)))

```

```{r preparation}
print(head(df_joined))

#past_0_29, past_30_59, past_60_89, past_90_119, past_120_149,past_150_plus -> set NA to 0
df_joined[is.na(df_joined)] = 0

# new column good or bad client
df_joined$raiting <- 0
#print(head(df_joined))

#df_joined %>% mutate(raiting = (df_joined$past_60_89+df_joined$past_90_119+df_joined$past_120_149+df_joined$past_150_plus))
#print(rowSums(df_joined[, c(past_60_89,past_90_119)]))
#df_joined$raiting <- (df_joined$past_60_89+df_joined$past_90_119+df_joined$past_120_149+df_joined$past_150_plus) / df_joined$total_months
df_joined$raiting <- as.numeric((df_joined$past_60_89+df_joined$past_90_119+df_joined$past_120_149+df_joined$past_150_plus))
print(df_joined$raiting)

write.csv(df_joined,"data/app_cc_record_join.csv", row.names = TRUE)
```