---
title: "ml_project_dataset_24months"
author: "Jeanette Lee, Kevin Stadelmann, Marc Weber, Patricia Wellhöfer"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    css: styles.css
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
---

# Introduction
[ACTION NEEDED: FILL IN]

Data Source: [link](https://www.kaggle.com/rikdifos/credit-card-approval-prediction)
(last update: 20-03-24)

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
library(magrittr)
library(data.table)
library(dplyr)
library(lubridate)
library(caret)
library(knitr)

opts_chunk$set(fig.align="center", echo=TRUE)

knit_hooks$set(inline=function(x) {
  prettyNum(x, big.mark=",")
})
```

# Data Import and Cleaning

```{r import-data, include=TRUE, cache=TRUE}
df.ar <- fread("data/application_record.csv", stringsAsFactors = TRUE)
df.cr <- fread("data/credit_record.csv", stringsAsFactors = TRUE)
```

## View Data
### Application Record Dataframe
DAYS_BIRTH and DAYS_EMPLOYED columns count backwards from "current" day (0), -1 means yesterday, etc.  
DAYS_EMPLOYED is positive, if the person is currently unemployed.
```{r view-df-ar, echo=FALSE}
kable(sapply(df.ar, class), col.names="type")
# [NOTE FOR RMD] Convert variables to factors: ID, FLAG_x, CNT_x; calculate age and work in years
head(df.ar) %>% kable()
# [NOTE FOR RMD] OCCUPATION_TYPE blank => replace with NA
```

### Credit Record Dataframe
MONTHS_BALANCE column counts backward from "current" month (0), -1 = previous month, -2 = 2 months prior, etc.  
STATUS column: 0 = 1-29 days past due, 1 = 30-59 days past due, 2 = 60-89 days overdue, 3 = 90-119 days overdue, 4 = 120-149 days overdue, 5 = Overdue or bad debts, write-offs for more than 150 days, C = paid off that month,  X = No loan for the month
```{r view-df-cr, echo=FALSE}
kable(sapply(df.cr, class), col.names="type")
# [NOTE FOR RMD] Convert ID to factor
head(df.cr) %>% kable()
# [NOTE FOR RMD] Add explanation for interpretation of MONTHS_BALANCE, STATUS columns
```

## Clean Data

### Cleaning application_record.csv
Remove duplicate IDs from application_record.csv
```{r clean-df-ar, collapse=TRUE, cache=TRUE}
df.ar$ID %>% unique() %>% length()

duplicates <- df.ar %>% group_by(ID) %>% summarise(count = n()) %>% filter(count > 1)
df.ar.mod <- df.ar %>% filter(!ID %in% duplicates$ID)

df.ar.mod$ID %>% unique() %>% length()
```

### Cleaning credit_record.csv
The original dataset was made available on 2020-03-24, so it is presumed current month is March 2020. 
From months_balance we create 2 new columns indicating the month and year for each observation. Observations older than 2 years are removed.

Each ID is assigned a "good" or "bad" customer status according to their credit history. If the customer paid off their balance in under 60 days (STATUS = 0, 1, or C) for more than 50% of their balance history, they are determined to be a "good" customer, otherwise they are determined to be a "bad" customer.
```{r clean-df-cr, include=TRUE, cache=TRUE}
# Extract month and year from months_balance in df.cr
df.cr$month_balance <- format(as.Date("2020-03-24") %m+% months(df.cr$MONTHS_BALANCE), "%B")
df.cr$year_balance <- format(as.Date("2020-03-24") %m+% months(df.cr$MONTHS_BALANCE), "%Y") %>% as.integer(.)

# Remove records older than 2 years
df.cr.mod <- df.cr %>% filter(MONTHS_BALANCE > -24)

# Assign customer status
df.status <- df.cr.mod %>%
  group_by(ID) %>%
  summarise(cnt_status_good = length(STATUS[STATUS==0|STATUS==1|STATUS=='C']), 
            cnt_status_bad = n() - cnt_status_good, 
            percent_good = round(cnt_status_good / n() * 100),
            percent_bad = round(cnt_status_bad / n() * 100)) %>%
  mutate(
    customer_status_good = ifelse(percent_good >= 50, 1, 0)
  ) %>%
  select(ID, customer_status_good)

df.status$customer_status_good <- as.factor(df.status$customer_status_good)
```

### Join datasets
The two cleaned datasets are joined and only records where IDs exist in both are kept.

Cleaning steps undertaken include: ... [ACTION NEEDED HERE]
```{r join-datasets, include=TRUE}

df.cc.raw <- inner_join(x = df.status, y = df.ar.mod, by = "ID") %>%
  mutate(ID = as.factor(ID),
         CNT_CHILDREN = as.factor(CNT_CHILDREN),
         CNT_FAM_MEMBERS = as.factor(CNT_FAM_MEMBERS),
         age_years = as.numeric(round(DAYS_BIRTH * (-1) / 365.25), 0),         # instead of DAYS_BIRTH
         work_years = as.numeric(round((DAYS_EMPLOYED * (-1) / 365.25), 0)),   # instead of DAYS_EMPLOYED, negative: Pensioner
         FLAG_MOBIL = as.factor(FLAG_MOBIL),
         FLAG_WORK_PHONE = as.factor(FLAG_WORK_PHONE),
         FLAG_PHONE = as.factor(FLAG_PHONE),
         FLAG_EMAIL = as.factor(FLAG_EMAIL))

# Change column names to lower case
setnames(df.cc.raw, tolower(names(df.cc.raw)))

# Look at the factor levels
df.cc.raw %>% 
  select(-id) %>% 
  select_if(is.factor) %>%
  sapply(., FUN = levels)

# flag_mobil = 1 for all IDs => exclude from final dataset

# occupation_type has "" level => replace with NA
levels(df.cc.raw$occupation_type)[levels(df.cc.raw$occupation_type)==""] <- "NA"
# df.cc.raw$occupation_type %>% table()                                        # check that "" are updated as NA

# Set order of levels in education type
df.cc.raw$name_education_type <- factor(df.cc.raw$name_education_type, levels=c("Lower secondary", "Secondary / secondary special", "Incomplete higher", "Higher education", "Academic degree"), ordered = TRUE)

# Convert -ve values to NA
df.cc.raw$work_years[df.cc.raw$work_years == '-1000'] <- NA

# Select columns
# months_balance, days_birth, days_employed and flag_mobil columns are excluded
df.cc <- df.cc.raw %>%
  select(id,
         code_gender,
         age_years,
         name_education_type,
         name_income_type,
         name_housing_type,
         name_family_status,
         cnt_children,
         cnt_fam_members,
         flag_own_car,
         flag_own_realty,
         work_years,
         amt_income_total,
         occupation_type,
         flag_work_phone,
         flag_phone,
         flag_email,
         customer_status_good)

```

### Check Final Dataframe
```{r summary-df-cc, include=TRUE, collapse=TRUE}
# Meta information from df.cc
nlevels(df.cc$id)

str(df.cc)

head(df.cc) %>% kable()
```